---
title: "100-histories"
output: html_document
date: "2024-05-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library("tidyverse")
```

Christian Denney:

Here is an updated set of data, along with figures after doing a little more QAQC etc. The attached .csv file contains:       
fish ID      
age of the fish (according to otolith aging)      
WY     
100 day cluster designations (1-4)     
150 day cluster designations (1-4)     
descriptive names of those clusters (HS = "high salinity", MS = "medium salinity", LS = "low salinity", FW = "freshwater")     
a selection column that indicates if the fish falls into your subsampling criteria (WY 2000, 2010, 2016 or 100 day cluster number 4)     
Additionally, I've included an updated figure showing all the clusters, as well as a version of Panel A with just the selected fish, and then finally, a panel plot of the 150 day clusters plotting calculated salinity instead of Sr isotope ratio.     

Let me know if you have any further questions!    


```{r}
df<-read_csv("Longfin smelt meta and tissues/all_fish_clust_ids.csv")

```

```{r}
 df %>% filter(WY %in% c("2000","2010","2016","2013")) %>% group_by(clust_name_100) %>% summarize(Count=n())
```


Mean strontium ratio up to day 100? Or mean pratical salinity up to day 100? 

```{r}
 df %>% filter(WY %in% c("2000","2010","2016","2013")) %>% filter(clust_name_100=="MS")
```