---
title: "202-kmer-gwas"
output: html_document
date: "2025-07-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

```{r}
library(tidyverse)
```

Need to set up a conda environment with:

python 2.7
R
R packages; MASS, Mvnpermute, matrixcalc

conda create -n kmers python=2.7
conda activate kmers
conda install R

mkdir kmers
cd kmers
wget https://github.com/voichek/kmersGWAS/releases/download/v0.3-beta/v0_3_beta.zip
unzip

Make kmers?

in outputs/201

_1_ need fastqs as a text file, 'LFS540.txt'

/home/maccamp/longfin-histories/data/alt-demultiplexed/LFS540_R1.fastq.gz
/home/maccamp/longfin-histories/data/alt-demultiplexed/LFS540_R2.fastq.gz

_2_ run kmc.  

```{sh, eval=FALSE}
/home/maccamp/kmers/external_programs/kmc_v3 -t2 -k31 -ci2 @LFS540.txt LFS540_kmc_canon ./ 1> kmc_canon.1 2> kmc_canon.2
 
/home/maccamp/kmers/external_programs/kmc_v3 -t2 -k31 -ci0 -b @LFS540.txt LFS540_kmc_all ./ 1> kmc_all.1 2> kmc_all.2

/home/maccamp/kmers/bin/kmers_add_strand_information -c LFS540_kmc_canon -n LFS540_kmc_all -k 31 -o LFS540-kmers_with_strand

## clean up
rm *.kmc*
```

Produces LFS540-kmers_with_strand

Should make each individual a directory
Can restrict to the set of fish that we run GWAS on