---
title: "206-polished-assembly"
output: html_document
date: "2025-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(snpR)
library(adegenet)
library(vcfR)
library(ggrepel)
```


/home/skieran/lfs/genome/lfs_run3_final/lfs_run3_scaffolds_final.fa

sym link to here:

/home/maccamp/genomes/longfin-polished/

```{sh, eval=FALSE}
module load bwa
srun -t 2:00:00 -p high --mem=16G --nodes=1 --ntasks-per-node=1 --cpus-per-task=2 bwa index lfs_run3_scaffolds_final.fa
```

Linking in data from: /home/maccamp/data/forage-fishes/longfin-previous
to /home/maccamp/longfin-histories/data/polish-saglam

```{sh, eval=FALSE}
bash ../../101.2-do-align.sh to-align.txt /home/maccamp/genomes/longfin-polished/lfs_run3_scaffolds_final.fa
```

Lots of stuff on the farm! Hope it works.


Trying YBAK


```{r}
library(tidyverse)
library(snpR)
library(adegenet)
library(vcfR)
library(ggrepel)
```

```{r}
yak<-read_csv("meta/yak.csv") %>% select(-Aligned, -Dedup, -Cov)
yak$Path<-gsub("saglam","polish-saglam",yak$Path)
yak %>% select(Path) %>% write_tsv("bamlists/yak-polish.bamlist", col_names = FALSE)
yak %>% select(SampleID) %>% write_tsv("bamlists/yak-polish.names", col_names = FALSE)
```

Call snps

```{sh, eval=FALSE}
srun -t 12:00:00 -p high --mem=32G --nodes=1 --ntasks-per-node=1 --cpus-per-task=12 $HOME/angsd/angsd -nthreads 12 \
-minInd 18 -bam bamlists/yak-polish.bamlist -ref /home/maccamp/genomes/longfin-polished/lfs_run3_scaffolds_final.fa \
-out outputs/206/snps-wgs  \
-minMaf 0.05 -minMapQ 20 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 2 -doPost 1 -postCutoff 0.9 -doPlink 2  > outputs/206/snps.out 2> outputs/206/snps.err &
```

        -> Number of sites retained after filtering: 62887 
   

```{sh, eval=FALSE}
#create vcf
plink --tped snps-wgs.tped --tfam snps-wgs.tfam  --out plink-binary --recode --allow-extra-chr --noweb
plink --ped plink-binary.ped --map plink-binary.map --recode vcf --allow-extra-chr -out plink

module load bcftools
bcftools reheader -s ../../bamlists/yak-polish.names -o plink-renamed.vcf plink.vcf
bcftools +fill-tags plink-renamed.vcf | bcftools view -i 'F_MISSING < 0.10' | bcftools view -i 'INFO/MAF > 0.01' > filtered.vcf
bcftools +prune -m 0.20 -w 50kb --nsites-per-win 1 --nsites-per-win-mode rand filtered.vcf > plink-pruned.vcf
```


Basic PCA

```{r}
snps<-import.snpR.data(genotypes = "outputs/206/plink-pruned.vcf", 
      sample.meta = yak)
```

```{r}
p <- plot_clusters(snps, facets = c("Pop"), viridis.option = "H")
p$plot$pca
```


Pretty much, that's it.


## LD

```{sh, eval=FALSE}
#long seqs are > 884174 bp
cat outputs/206/long-seqs.tsv  | while read line; do bcftools view -Ov -r $line outputs/206/filtered.vcf.gz > outputs/206/vcfs/$line.vcf; done;

for f in *.vcf; do plink --vcf $f --r2 inter-chr --ld-window-r2 0.1 --out `basename $f vcf`ldf --allow-extra-chr --double-id; done;

```

```{r, eval=FALSE}
files<-list.files("outputs/206/vcfs",pattern = "*.ldf.ld", full.names = TRUE)

plotLd<-function(file) {
  chrom<-gsub("outputs/206/vcfs/","",file)
  chrom<-gsub(".ldf.ld","", chrom)
  lc<-read.delim(file,sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% arrange(R2) %>%  filter(R2 >0.5)

  ggplot(lc) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle(paste0(chrom))+
  theme_bw() +
  theme(panel.grid = element_blank())
ggsave(paste0("outputs/206/chrom-ld/",chrom,".pdf"))
}

lapply(files, plotLd)
```


Scaffold 26 is interesting

```{r}
lc<-read.delim("outputs/206/vcfs/lcl|scaffold_26.ldf.ld",sep="",stringsAsFactors=FALSE) %>% as_tibble() %>% arrange(R2) %>%  filter(R2 >0.2)

  ggplot(lc) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle("lcl|scaffold_26 LD")+
  theme_bw() +
  theme(panel.grid = element_blank()) 
  
ggsave("outputs/206/scaffold26-ld-plot.pdf")
```

````{r}
ggplot(lc) +
  geom_point(aes(x=BP_A, y=BP_B, color=R2), alpha=0.5) +
  scale_color_gradient(low="khaki1", high="red") + 
  ggtitle("lcl|scaffold_26 LD")+
  theme_bw() +
  theme(panel.grid = element_blank()) +
  xlim(8.25e6, max(lc$BP_A))
```

```{sh, eval=FALSE}
bcftools view -Ov -r 'lcl|scaffold_26:8250000-' outputs/206/filtered.vcf.gz > outputs/206/lcl26-sub.vcf
```


Chrom PCS

```{r, eval=FALSE}
files<-list.files("outputs/206/vcfs",pattern = "*.vcf", full.names = TRUE)

plotPCA<-function(file) {
  meta<-yak
  chrom<-gsub("outputs/206/vcfs/","",file)
  chrom<-gsub(".vcf","", chrom)
  
  vcf<-read.vcfR(file=file)
  genind<-vcfR2genind(vcf)
  genind@pop<-as.factor(meta$Pop)

gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)

df<-pca1$li
df<-df %>% bind_cols(meta)
eig<-pca1$eig/sum(pca1$eig)*100


pc12<-ggplot(df) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=Pop),pch=21, alpha=0.75, cex=2) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "top") +
  geom_text_repel(aes(x=Axis1, y=Axis2,label=SampleID))
pc12

ggsave(paste0("outputs/206/chrom-pcs/",chrom,".pdf"))
}

lapply(files, plotPCA)
```


## Check loadings
```{r}
vcf<-read.vcfR(file="outputs/206/lcl26-sub.vcf")
genind<-vcfR2genind(vcf)
genind@pop<-as.factor(yak$Pop)
```


```{r}
gen<-genind
X <- tab(gen, NA.method="mean")

pca1 <- dudi.pca(X,scannf=FALSE,scale=FALSE,nf=3)

df<-pca1$li
df<-df %>% bind_cols(yak)

eig<-pca1$eig/sum(pca1$eig)*100
```

```{r}
pc12<-ggplot(df) + 
  geom_point(aes(x=Axis1, y=Axis2, fill=Pop), alpha=0.75, cex=2, pch=21) +
  theme_bw() +
  xlab(paste0("PC1 ",round(eig[1],2), "%")) +
  ylab(paste0("PC2 ",round(eig[2],2), "%")) +
  theme(panel.grid = element_blank()) +
  scale_fill_viridis_d(option="H") +
  guides(fill = guide_legend(override.aes = list(pch = 22))) +
  theme(legend.position = "top") +
  geom_text_repel(aes(x=Axis1, y=Axis2,label=SampleID))
pc12
```

```{r}
loadings<-pca1$c1 %>% as_tibble()
loadings$Allele<-rownames(pca1$c1) 

loadings
loadings$Position<-gsub("lcl\\|.*_|\\.\\d$","",loadings$Allele)
loadings$Position<-as.numeric(loadings$Position)
loadings$Chrom<-gsub("_\\d+\\.[0|1]","",loadings$Allele)
tops<-loadings %>% arrange(-CS1) %>% slice_max(order_by = CS1,prop = .01) %>% select(-CS2, -CS3) %>% 
  mutate(MajorMinor=gsub("lcl\\|*\\.","",Allele)) 

tops %>% relocate(Chrom, Position, CS1)

```

lcl|scaffold_26	8445555	lcl|scaffold_26_8445555	T	C	.	.	PR;F_MISSING=0;NS=20;AN=40;AF=0.325;MAF=0.325;AC=13;AC_Het=13;AC_Hom=0;AC_Hemi=0;HWE=0.0575543;ExcHet=0.0527742	GT	0/1	0/0	0/1	0/1	0/1	0/0	0/0	0/1	0/1	0/0	0/1	0/1	0/1	0/1	0/0	0/1	0/1	0/0	0/0	0/1
