---
title: "215-redo-rad"
output: html_document
date: "2025-11-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
library(RColorBrewer)
library(vcfR)
library(adegenet)
library(tidyverse)
```

The goal is to go to the radseq from the G3 paper and realign etc.

(base) ➜  github git clone https://github.com/MacCampbell/delta-smelt/
Cloning into 'delta-smelt'...


```{r}
lh<-read_csv("../delta-smelt/metadata/2012-all-samples.csv")
lh
```

We want to move all the .fastqs to my longfin-histories/redo Looks like they are gzipped

(base) maccamp@farm:~/longfin-histories/data/redo$ cut -f 4 -d ',' ~/delta-smelt/metadata/2012-all-samples.csv  | while read line; do cp ~/ds-lh/data/$line*fastq ./; done;

/group/millermrgrp2/shannon/projects/DS_history/data/BMAG048/ACTGAT$ 

Files are in 
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG043/ATCACG/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG043/CGATGT/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG044/ACAGTG/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG044/GCCAAT/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG045/ACTTGA/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG045/CAGATC/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG045/GATCAG/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG046/AGTCAA/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG046/AGTTCC/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG046/CTTGTA/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG047/ATGTCA/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG047/CCGTCC/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG047/GTCCGC/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG047/GTGAAA/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG048/ACTGAT/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG048/ATTCCT/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG048/CGTACG/
/group/millermrgrp2/shannon/projects/DS_history/data/BMAG048/GTGGCC/

cat dirs  | while read line; do ln -s $line*.fastq.gz .; done;
 
 
we want a file like
Ht20-33_2012_A05_R1.fastq.gz    Ht20-33_2012_A05_R2.fastq.gz    Ht20-33_2012_A05

```{r}
lh %>% select(`Sequence File Name`) %>% mutate(Forward=paste0(`Sequence File Name`,"_R1.fastq.gz")) %>% mutate(Reverse=paste0(`Sequence File Name`,"_R2.fastq.gz")) %>% relocate(Forward, Reverse, `Sequence File Name`) %>%
  write_tsv("215.1-to-align.tsv", col_names = FALSE)
```

165 samples to align

